# ECON HISTORY - ASSIGNMENT 2

### Benramdane, Blanchon, Garson, Maestri

## 0. Importing libraries and inspectioning data

```{r}
#install.packages("tidyverse")
#install.packages("skimr")
#install.packages("stargazer")
#install.packages("sandwich")
#install.packages("aod")
library(tidyverse)
library(skimr)
library(stargazer)
library(sandwich)
library(aod)
library(countrycode)
library(lpirfs)
jst_data <- read.csv('./JSTdatasetR6.csv')
names(jst_data)[0:5]
```

## 1. Estimate first logit model

### a. Select and refine data for logit model

```{r}
logit_data <- jst_data %>%
  arrange(country, year) %>%
  mutate(tloans_real = tloans/cpi, 
         ln_tloans_real = log(tloans_real),
         year = as.integer(year),
         .after = tloans) %>%
  mutate(ln_diff_tloans = ln_tloans_real - lag(ln_tloans_real, n=1),
         .after = ln_tloans_real) %>%
  select(country, year, crisisJST,
         tloans_real, ln_tloans_real, 
         ln_diff_tloans, tloans, cpi,
         gdp)
```

#### b. Running the logit model and exporting results

```{r}
#Compute model
logit_credit <- glm(crisisJST ~ lag(ln_diff_tloans, 1) +
                      lag(ln_diff_tloans, 2) + 
                      lag(ln_diff_tloans, 3) + 
                      lag(ln_diff_tloans, 4) + 
                      lag(ln_diff_tloans, 5) +
                      factor(country),
                    family = binomial(link = "logit"),
                    data = logit_data)

#Plot results
stargazer(logit_credit,
          type = "text",
          align = TRUE,
          omit = c("country"))
```

#### c. Test for joint significance of lagged variables

```{r}
# Perform Wald test
wald.test(
  b = coef(logit_credit),             # Coefficients from the full model
  Sigma = vcov(logit_credit),         # Variance-covariance matrix of the coefficients
  Terms = 2:6                       # Indices of lagged predictors in the model
)
```

#### d. Import BVX crisis list and merge with JST data

```{r}
#Use countrycodes to merge
jst_data$countrycode <- countrycode(jst_data$country, 
                                    origin = "country.name", 
                                    destination = "iso3c")

#Read BVX data
bvx_crises <- read.csv("./bvx_crises_list.csv") %>%
  select(country, year, bvx_crisis)
bvx_logit_data <- left_join(logit_data, bvx_crises, by = join_by(country, year))
```

#### e. Estimate logit model for BVX crises

```{r}
#Compute model
bvx_logit_credit <- glm(bvx_crisis ~ lag(ln_diff_tloans, 1) +
                      lag(ln_diff_tloans, 2) + 
                      lag(ln_diff_tloans, 3) + 
                      lag(ln_diff_tloans, 4) + 
                      lag(ln_diff_tloans, 5) +
                        factor(country),
                    family = binomial(link = "logit"),
                    data = bvx_logit_data)

#Plot results
stargazer(bvx_logit_credit,
          type = "text",
          align = TRUE,
          omit = c("country"))
```

#### f. Test for joint significance of lagged variables in BVX crises subset

```{r}
# Perform Wald test
wald.test(
  b = coef(bvx_logit_credit),             # Coefficients from the full model
  Sigma = vcov(bvx_logit_credit),         # Variance-covariance matrix of the coefficients
  Terms = 2:6                       # Indices of lagged predictors in the model
)
```

#### h. Use 5-year change in credit-to-GDP ratio as a predictor

```{r}
# Construct 5-year-change indicator
logit_data <- logit_data %>%
  mutate(credit_to_gdp = tloans / gdp) %>%
  mutate(credit_to_gdp_fyc_pp = (credit_to_gdp - lag(credit_to_gdp, n=5))*100) 

# Estimate new logit model
#Compute model
logit_credit_to_gdp <- glm(crisisJST ~ credit_to_gdp_fyc_pp + 
                      factor(country),
                    family = binomial(link = "logit"),
                    data = logit_data)

#Plot results
stargazer(logit_credit_to_gdp,
          type = "text",
          align = TRUE,
          omit = c("country"))
```

## Local Projections

#### a. Load and prepare data for analysis  

```{r}
# Clear environment
rm(list = ls())

# Load datasets and merge 
jst_data <- read.csv('./JSTdatasetR6.csv')
rec_dummies <- read.csv("./RecessionDummies.csv")

# Merge on iso code and year
lp_data <- left_join(jst_data, rec_dummies, by = join_by(iso, year)) %>%
  select(year, iso, country,
         gdp, rgdpbarro, 
         N, F)

# Define the dependent variable: real GDP per capita growth
lp_data <- lp_data %>%
  group_by(iso) %>%
  mutate(d_y = log(rgdpbarro) - lag(log(rgdpbarro))) %>%
  ungroup()

# Replace NAs with zeros to avoid problems with lps
# lp_data$N <-replace_na(lp_data$N, 0)
# lp_data$F <-replace_na(lp_data$F, 0)
# Filter out missing values
# lp_data <- lp_data %>%
#   filter(!is.na(N))
```

#### b. Estimate the path of real GDP per capita after normal recessions

```{r}
# Run local projections
lp_normal <- lp_lin_panel(
  data_set = lp_data,
  endog_data = "d_y",        # Dependent variable
  cumul_mult = TRUE,         # Cumulative multipliers
  shock = "N",               # Normal recession
  diff_shock = FALSE,        # No differencing of shocks
  panel_model = "within",    # Fixed effects model
  confint = 1.96,            # 95% confidence interval
  hor = 5                    # Horizons: 1 to 5 years
)

# Inspect results
# print(lp_normal)

# Plot impulse response functions (IRFs)
plot(lp_normal)
```

#### c. Estimate the path of real GDP per capita after financial recessions

```{r}
# Run local projections
lp_financial <- lp_lin_panel(
  data_set = lp_data,
  endog_data = "d_y",        # Dependent variable
  cumul_mult = TRUE,         # Cumulative multipliers
  shock = "F",               # Financial recession
  diff_shock = FALSE,        # No differencing of shocks
  panel_model = "within",    # Fixed effects model
  confint = 1.96,            # 95% confidence interval
  hor = 5                    # Horizons: 1 to 5 years
)

# Inspect results
# print(lp_normal)

# Plot impulse response functions (IRFs)
plot(lp_financial)
```
