---
---
---

# Econ History Assignment 2

Ali BENRAMDANE, Mathilde BLANCHON, Jonathan GARSON, Andrea MAESTRI

## Loading the library

```{r, results='hide', warning=FALSE, message = FALSE}
library(data.table)
library(fixest)
library(stargazer)
library(aod)
library(readxl)
library(countrycode)
library(pROC)
library(lpirfs)
```

## Data

```{r}
jst = fread("JSTdatasetR6.csv")
rec = fread("RecessionDummies.csv")
bvx = fread("bvx_crises_list.csv")
```

## Part A

### Logit

#### JST Crisis

Estimate a logit model with country fixed effects and with five lags of log changes in real credit as a predictor for JST crises and document the results in table.

-   Test for joint significance of five lags credit growth.

-   Repeat with BVX crisis date

-   Used 5-year change in the ratio of credit over GDP predictor

    ```{r, warning=FALSE}
    jst[, tloans_real := tloans/cpi]
    jst[, ln_tloans_real := log(tloans_real),]
    jst[, ln_diff_tloans := ln_tloans_real - shift(ln_tloans_real, 1, type = "lag"), by = country]

    logit_credit = glm(crisisJST ~ 
          shift(ln_diff_tloans, 1, type = "lag") 
          + shift(ln_diff_tloans, 2, type = "lag") 
          + shift(ln_diff_tloans, 3, type = "lag") 
          + shift(ln_diff_tloans, 4, type = "lag") 
          + shift(ln_diff_tloans, 5, type = "lag") 
          + factor(country)
        ,family = binomial(link = "logit"), 
        data = jst)

    stargazer(logit_credit,
              type = "text",
              align = TRUE,
              omit = c("country"))
    ```

    We know test for the joint significance of the lag variables

    ```{r}
    wald.test(
      b = coef(logit_credit),  # Coefficients from the full model
      Sigma = vcov(logit_credit),  # Variance-covariance matrix of coefficients
      Terms = 2:6 # Indices of lagged predictors in the model
    )
    ```

#### BVX Crisis

```{r, warning=FALSE}
# We merge 
setnames(bvx, "countrycode", "iso")
bvx = bvx[, country := NULL]
jst = merge(jst, bvx, all.x = TRUE, by = c("iso", "year")) 

# We regress

logit_bvx = glm(bvx_crisis ~ 
      shift(ln_diff_tloans, 1, type = "lag") 
      + shift(ln_diff_tloans, 2, type = "lag") 
      + shift(ln_diff_tloans, 3, type = "lag") 
      + shift(ln_diff_tloans, 4, type = "lag") 
      + shift(ln_diff_tloans, 5, type = "lag") 
      + factor(country)
    ,family = binomial(link = "logit"), 
    data = jst)

stargazer(logit_credit,
          type = "text",
          align = TRUE,
          omit = c("country"))
```

```{r}
wald.test(
  b = coef(logit_bvx),  # Coefficients from the full model
  Sigma = vcov(logit_bvx),  # Variance-covariance matrix of coefficients
  Terms = 2:6 # Indices of lagged predictors in the model
)
```

#### GDP ratio

Used 5-year change in the ratio of credit over GDP as a predictor

```{r, warning = FALSE}
jst[, `:=` (gdp_real = gdp/cpi), by = c("country")]
jst[, creditgdp := tloans_real/gdp_real, by = c("country")]
jst[, creditgdpchange := creditgdp - shift(creditgdp, 5, type = "lag")]

logit_gdp = glm(crisisJST ~ 
                creditgdpchange 
                + factor(country)
                ,family = binomial(link = "logit"), 
                data = jst)

stargazer(logit_gdp,
          type = "text",
          align = TRUE,
          omit = c("country"))
```

### Model estimation

#### ROC test

Compare in-sample and out of sample ROC, estimating the model until 1984 and predicting crisis for post-1984 years.

```{r, warning=FALSE}
# we split data in two
pre1984 = jst[year <= 1984,]
post1984 = jst[year > 1984,]

# we test on our pre-1984 data
logit_1984 = glm(crisisJST ~ 
      shift(ln_diff_tloans, 1, type = "lag") 
      + shift(ln_diff_tloans, 2, type = "lag") 
      + shift(ln_diff_tloans, 3, type = "lag") 
      + shift(ln_diff_tloans, 4, type = "lag") 
      + shift(ln_diff_tloans, 5, type = "lag") 
      + country
    ,family = binomial(link = "logit"), 
    data = pre1984)

# Predict probabilities for training data (in-sample)
pre1984$predicted_prob <- predict(logit_1984, pre1984, type = "response")

# Predict probabilities for testing data (out-of-sample)
post1984$predicted_prob <- predict(logit_1984, post1984, type = "response")

# Compute ROC curve for in-sample data
roc_in_sample <- roc(pre1984$crisisJST, pre1984$predicted_prob)

# Compute ROC curve for out-of-sample data
roc_out_sample <- roc(post1984$crisisJST, post1984$predicted_prob)

# Plot ROC curves
plot(roc_in_sample, col = "blue", main = "In-Sample vs. Out-of-Sample ROC", legacy.axes = TRUE)
plot(roc_out_sample, col = "red", add = TRUE)

# Add legend
legend("bottomright", legend = c("In-Sample", "Out-of-Sample"),
       col = c("blue", "red"), lty = 1, cex = 0.8)

# Print AUC values
cat("In-Sample AUC: ", auc(roc_in_sample), "\n")
cat("Out-of-Sample AUC: ", auc(roc_out_sample), "\n")

```

#### Money as predictor

Compare the baseline model to a logit model with money as a predictor.

```{r, warning=FALSE}
jst[, money_real := money/cpi]
jst[, ln_money_real := log(money_real),]
jst[, ln_diff_money_real := ln_money_real - shift(ln_money_real, 1, type = "lag"), by = country]

logit_money = glm(crisisJST ~ 
      shift(ln_diff_money_real, 1, type = "lag") 
      + shift(ln_diff_money_real, 2, type = "lag") 
      + shift(ln_diff_money_real, 3, type = "lag") 
      + shift(ln_diff_money_real, 4, type = "lag") 
      + shift(ln_diff_money_real, 5, type = "lag") 
      + factor(country)
    ,family = binomial(link = "logit"), 
    data = jst)

stargazer(logit_money,
          type = "text",
          align = TRUE,
          omit = c("country"))
```

#### Public debt as predictor

```{r, warning=FALSE}
jst[, diff_debtgdp := debtgdp - shift(debtgdp, 1, type = "lag"), by = c("country")]

logit_gdp = glm(crisisJST ~ 
      shift(diff_debtgdp, 1, type = "lag") 
      + shift(diff_debtgdp, 2, type = "lag") 
      + shift(diff_debtgdp, 3, type = "lag") 
      + shift(diff_debtgdp, 4, type = "lag") 
      + shift(diff_debtgdp, 5, type = "lag") 
      + factor(country)
    ,family = binomial(link = "logit"), 
    data = jst)

stargazer(logit_gdp,
          type = "text",
          align = TRUE,
          omit = c("country"))
```

## Part B

```{r}
jst = merge(jst, rec, by = c("year", "iso"))
jst[, d_y := log(rgdpbarro) - shift(log(rgdpbarro), 1, type = "lag")]
```

### Normal Recessions

```{r}
lp_normal <- lp_lin_panel(
  data_set = jst,
  endog_data = "d_y",        # Dependent variable
  cumul_mult = FALSE,         # Cumulative multipliers
  shock = "N",               # Normal recession
  diff_shock = FALSE,        # No differencing of shocks
  panel_model = "within",    # Fixed effects model
  confint = 1.96,            # 95% confidence interval
  hor = 5                    # Horizons: 1 to 5 years
)

# Plot impulse response functions (IRFs)
plot(lp_normal)
```

### Financial Recessions

```{r}
lp_normal <- lp_lin_panel(
  data_set = jst,
  endog_data = "d_y",        # Dependent variable
  cumul_mult = FALSE,         # Cumulative multipliers
  shock = "F",               # Normal recession
  diff_shock = FALSE,        # No differencing of shocks
  panel_model = "within",    # Fixed effects model
  confint = 1.96,            # 95% confidence interval
  hor = 5                    # Horizons: 1 to 5 years
)

# Plot impulse response functions (IRFs)
plot(lp_normal)
```
